Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-08-31 draft.tex
modify lemmas 1 and 2 for homogenized delta

Future payoffs are discounted with parameters $\delta_i$, where 1/$\delta_i$ is player $i$'s discount rate for $i \in (g, s, c, p)$, $0\leq\delta_i\leq1$. Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \delta_i^{n-1}$.

@@

\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta_i} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta_i} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}

\item \textit{$\frac{-\alpha}{1-\de_p} \geq \frac {\beta}{1-\delta_c}$: reunification is more important for the patron to avoid than for the international community to achieve.}

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{(1-\delta_s)}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{(1-\delta_i)}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, $L_s$ forever with probability $p_2$, and the status quo payoff forever with probability $1-p_1-p_2$. Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta_i}$.

In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2). The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= \alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

Thus the difference in continuation values from switching from the status quo to reunification for player $p$ is $\frac{\alpha + R_p^{SQ} 
}{1 -\de_p}-R_p^{RU}$. For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de_c} -R_c^{RU}$. Neither player will be willing to invest in its least preferred state. That is, $R_p^{RU} = R_c^{REC} =0$. So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de_c}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ} 
\right)}{1 -\de_p}$ each period toward maintaining it.

Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g. counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de_c} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de_p} > \frac{\beta+ R_c^{SQ}}{1 -\de_c}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\frac{-\alpha}{1 -\de_p}  > \frac{\beta}{1 -\de_c}$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

>>>>

Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates. This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \de^{n-1}$.

@@

\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}

\item \textit{$-\alpha \geq \beta$: reunification is more important for the patron to avoid than for the international community to achieve.}

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, $L_s$ forever with probability $p_2$, and the status quo payoff forever with probability $1-p_1-p_2$. Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta}$.

In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2). The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= \alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

Thus the difference in continuation values from switching from the status quo to reunification for player $p$ is $\frac{\alpha + R_p^{SQ} 
}{1 -\de}-R_p^{RU}$. For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de} -R_c^{RU}$. Neither player will be willing to invest in its least preferred state. That is, $R_p^{RU} = R_c^{REC} =0$. So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ} 
\right)}{1 -\de}$ each period toward maintaining it.

Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g. counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $-\alpha  > \beta$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
